---
title: "2024 NH3-loss Statistical Models, outdoor only - cube root"
author: "Maarit Mäenpää + NJH"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document:
    reference_docx: landscape.docx
  pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data, include=FALSE}

# Packages
library(tidyr)
library(plyr)
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)
library(emmeans)
library(interactions)
library(ggpubr)
library(car)
library(glmmTMB)
library(DHARMa)
library(tidyverse)

#
#
#

# Data
load(here::here("data/Wrangled/NH3predict.Rda"))

# Remove potential NAs from base-model variables
NH3predict <- NH3predict %>% drop_na(c(Fertiliser, Application, Method))

# Subset to only outdoor data
NH3predict <- filter(NH3predict, Place=="Outdoor")

NH3predict <- NH3predict[which(is.na(NH3predict$NH3loss)==FALSE &
                                 is.na(NH3predict$Fertiliser)==FALSE & 
                                 is.na(NH3predict$Application)==FALSE &
                                 is.na(NH3predict$Method)==FALSE &
                                 is.na(NH3predict$Ref)==FALSE),]

# Removing a study with large negative values for NH3loss
NH3predict <- NH3predict[which(NH3predict$Ref!="Velthof et al. 1990"),]

#
#
# 

## DEFINING VARIABLES


# Defining class of vectors in the analysis

#
# Model 1 (A1)

NH3predict$Fertiliser <- factor(NH3predict$Fertiliser)
NH3predict$Place <- factor(NH3predict$Place)
NH3predict$Method <- factor(NH3predict$Method)
NH3predict$Application <- factor(NH3predict$Application)
NH3predict$Ref <- factor(NH3predict$Ref)

#
# Model 2 (A2)
NH3predict$Application.rate <- as.numeric(NH3predict$Application.rate)

#
# Model 3 (A3)
NH3predict$SoilpH <- as.numeric(NH3predict$SoilpH)
summary(NH3predict$SoilpH)
NH3predict[which(NH3predict$SoilpH>10),]

# pH categorisation
NH3predict$pH_cat <- factor(ifelse(NH3predict$SoilpH<=7, "normal", "chalky"))
ddply(NH3predict, c("pH_cat"), summarise, N=length(NH3loss))

ggplot(NH3predict, aes(x=SoilpH, fill=pH_cat)) + geom_histogram(binwidth=0.1, alpha=.5, position="identity")
ggplot(NH3predict, aes(x=SoilpH, fill=pH_cat)) + geom_density(alpha=.3)

#
# Model 4 (A4)
NH3predict$Clay <- as.numeric(as.character(NH3predict$Clay))

#
# Model 5 (A5)
NH3predict$SOC <- as.numeric(as.character(NH3predict$SOC))
# SOC categorisation 
# Organic soils: >30% organic matter
# >12% SOC
NH3predict$SOC_cat <- factor(ifelse(NH3predict$SOC>=12, "organic", "mineral"))
ggplot(NH3predict, aes(x=SOC, fill=SOC_cat)) + geom_histogram(binwidth=0.1, alpha=.5, position="identity")
ggplot(NH3predict, aes(x=SOC, fill=SOC_cat)) + geom_density(alpha=.3)

#
# Model 6
NH3predict$Crop <- factor(NH3predict$Crop)
# Cover crop categorisation
NH3predict$Cover <- NH3predict$Crop
levels(NH3predict$Cover)[levels(NH3predict$Cover)=="alfalfa"] <- "Arable" 
levels(NH3predict$Cover)[levels(NH3predict$Cover)=="cereal"] <- "Arable"
levels(NH3predict$Cover)[levels(NH3predict$Cover)=="grass"] <- "Grass"
levels(NH3predict$Cover)[levels(NH3predict$Cover)=="beans"] <- "Arable"
levels(NH3predict$Cover)[levels(NH3predict$Cover)=="maize"] <- "Arable"
levels(NH3predict$Cover)[levels(NH3predict$Cover)=="none"] <- "None"
levels(NH3predict$Cover)[levels(NH3predict$Cover)=="other"] <- "Other"
levels(NH3predict$Cover)[levels(NH3predict$Cover)=="trees"] <- "Other"
summary(NH3predict$Cover)

#
# Model 7-8
NH3predict$Rainfall <- as.numeric(as.character(NH3predict$Rainfall))
NH3predict$AirTemperature <- as.numeric(as.character(NH3predict$AirTemperature))
NH3predict$Duration <- as.numeric(as.character(NH3predict$Duration))
NH3predict$Rain.mean <- NH3predict$Rainfall/NH3predict$Duration

# Model 9
# Binarising irrigation
NH3predict$irrigation_cat <- factor(ifelse(NH3predict$irrigation=="0", 0, 1))

# Model 10
NH3predict$WaterPerDay <- as.numeric(as.character(NH3predict$WaterPerDay))

#
#
# Reordering factor levels

NH3predict$Fertiliser <- factor(NH3predict$Fertiliser, levels = c("urea+", "UAN", "ammonium+1", "ammonium+2"))
NH3predict$Method <- relevel(NH3predict$Method, "Micromet")
NH3predict$Application <- relevel(NH3predict$Application, "Broadcast") 
NH3predict$Place <- relevel(NH3predict$Place, "Outdoor") 
NH3predict$pH_cat <- relevel(NH3predict$pH_cat, "normal") 

# Model response variable
NH3predict$NH3loss <- as.numeric(as.character(NH3predict$NH3loss))
```

## Model A_base

To reduce the number of parameters in the statistical model, we categorised different fertilisers into categories consisting of;

**Urea**: Fertilisers where the main component was urea

**UAN**: UAN

**ammonium+1**: Fertilisers including a low percentage of ammonium

**ammonium+2**: Fertilisers with higher percentage of ammonium

The table below depicts the categorisation of fertilisers, and the number of observations in each sub-category.

```{r Fertiliser, include=TRUE, echo=FALSE}
grouped<-NH3predict[which(NH3predict$Application=="Broadcast" | NH3predict$Application=="Liquid"),]
grouped %>% group_by(FertiliserType) %>% summarise(count = n(),mean_emm = mean(NH3loss),)

ddply(NH3predict, c("Fertiliser", "Fertiliser"), summarise, N=length(NH3loss))

```

In Model A_base, we examined the effects of application method (Application) and measurement method (Method) on NH3 emissions. The tables below show the number of observations within each category examined.

```{r include=TRUE, echo=FALSE}
ddply(NH3predict, c("Application"), summarise, N=length(NH3loss))

ddply(NH3predict, c("Method"), summarise, N=length(NH3loss))


```

Table below shows counts of all categories of variables:

```{r include=TRUE, echo=FALSE}

table(NH3predict$Fertiliser, NH3predict$Application, NH3predict$Method)

```

We analysed the effects of Fertiliser types (Urea, UAN, ammonium+1, and ammonium+2), Application method (Broadcast, Liquid, Incorporated, and Injected), the measurement method (Micromet, 15N, Closed chamber, Drager-Tube, Semi-open, Ventilated chambers, and Windtunnel), and whether the study was conducted in the field or in the laboratory, using a general linear mixed effects model (lmer, R-package lme4, Bates et al. 2015). Because the observations originated from different studies, each study most often contributing more than one observation, the reference for study was used as a random effect, to account for non-independence due to repeated measurements.

We used NH3-loss (%), calculated as relative difference between ammonia emissions from a test site and a control site proportional to the control, as a response variable. There were `r length(NH3predict[which(NH3predict$NH3loss<0),]$NH3loss)` negative observations out of the total `r length(NH3predict$NH3loss)`.

The frequency distribution of the untransformed NH3-loss measurements is shown below.

```{r include=TRUE, echo=FALSE}

hist(NH3predict$NH3loss, breaks=50)

```

To account for non-normality, the response variable was transformed with a square root transformation. We applied a constant of 4 to allow the analysis to include negative values.

To account for heteroscedasticity, we used weighted ordinary least squares (WOLS, Olvera Astivia & Bruno 2019). We built a vector giving larger weight to observations with smaller variation. To this end, we extracted absolute residuals and fitted values from an unweighted version of Model A_Base (see below for code). We then applied a regression model between the absolute residuals (response variable and fitted values (predictor variable), to estimate the relationship between the amount of variation and the data and the model estimates. To attain a vector with higher values attached to lower variation, we used the inverse of the relationship estimated in the regression model. The weights correspond to increased weight with decrease in variation, thus giving more precedence to more accurate data.

The R-code for the acquiring the weights, and the final model is shown below.

```{r, include=TRUE, echo=FALSE}
## MODEL A_Base ¨

# Model for calculating weights
model <- lmer(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Fertiliser + Application + Method  + (1|Ref),
              control=lmerControl(optimizer="bobyqa"),
              data=NH3predict)

# Calculating weights
NH3predict$Weight <- 1/lm(abs(resid(model)) ~ fitted(model))$fitted.values^2

# Model A_Base
m1_lmer_p1 <- lmer(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Fertiliser + Application + Method  + (1|Ref),
                   weights=Weight,
                   control=lmerControl(optimizer="bobyqa"),
                   data=NH3predict)

```

The model assumptions are checked in the figures below:

```{r include=TRUE, echo=FALSE}
#
# Model assumptions
par(mfrow=c(1,3))
plot(fitted(m1_lmer_p1), NH3predict$Weight*resid(m1_lmer_p1)) 
abline(h=0) # OK
qqnorm(resid(m1_lmer_p1))
qqline(resid(m1_lmer_p1)) 
hist(resid(m1_lmer_p1)) 
par(mfrow=c(1,1))


```

NH3-loss was different for different fertilisers (F(`r round(anova(m1_lmer_p1)$NumDF[1], 0)`, `r round(anova(m1_lmer_p1)$DenDF[1],0)`) =`r round(anova(m1_lmer_p1)$F[1], 2)`, p `r ifelse(round(anova(m1_lmer_p1)[[6]][1], 3)=="0", "<0.001", paste("=",round(anova(m1_lmer_p1)[[6]][1], 3)))`), as were the emissions from different application methods (F(`r round(anova(m1_lmer_p1)$NumDF[2], 0)`, `r round(anova(m1_lmer_p1)$DenDF[2],0)`) =`r round(anova(m1_lmer_p1)$F[2], 2)`, p `r ifelse(round(anova(m1_lmer_p1)[[6]][2], 3)=="0", "<0.001", paste("=",round(anova(m1_lmer_p1)[[6]][2], 3)))`). Laboratory measurements were significantly different from the measurements taken from the field (F(`r round(anova(m1_lmer_p1)$NumDF[4], 0)`, `r round(anova(m1_lmer_p1)$DenDF[4],0)`) =`r round(anova(m1_lmer_p1)$F[4], 2)`, p `r ifelse(round(anova(m1_lmer_p1)[[6]][4], 3)=="0", "<0.001", paste("=",round(anova(m1_lmer_p1)[[6]][4], 3)))`). Overall, the measurement method did not explain variation in NH3 emissions (F(`r round(anova(m1_lmer_p1)$NumDF[3], 0)`, `r round(anova(m1_lmer_p1)$DenDF[3],0)`) =`r round(anova(m1_lmer_p1)$F[3], 2)`, p `r ifelse(round(anova(m1_lmer_p1)[[6]][3], 3)=="0", "<0.001", paste("=",round(anova(m1_lmer_p1)[[6]][3], 3)))`).

```{r, include=FALSE, echo=FALSE}

anova(m1_lmer_p1)

```

The pairwise-comparisons of factor-level effects can be seen below. The comparisons are made to urea (in Fertiliser), Broadcast (in Application) and Micromet (in Method). The estimates are given in the sqrt(x+4) -scale.

\small

```{r include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

# Results
summary(m1_lmer_p1)

```

\normalsize

The effect sizes (marginal means) for each category of each factor are presented below, parametrised to urea (in Fertiliser), Broadcast (in Application), and Micromet (in Method). The effect sizes are given at the back-transformed scale (NH3-loss %).

\small

```{r include=TRUE, echo=FALSE}

#
##### 
# Estimates for each main effect

# Inverse function
inv_signed_cube_root <- function(y) sign(y) * abs(y)^3

#
# Fertiliser

# Effect sizes
at <- list("Application"="Broadcast", "Method"="Micromet")
m1_est_F <- emmeans(m1_lmer_p1, ~ Fertiliser, at=at)
m1_est_F  <- data.frame(m1_est_F) 

m1_est_F$emmean <- as.numeric(as.character(m1_est_F$emmean))
m1_est_F$emmean <- inv_signed_cube_root(m1_est_F$emmean)
m1_est_F$lower.CL <- inv_signed_cube_root(m1_est_F$lower.CL)
m1_est_F$upper.CL <- inv_signed_cube_root(m1_est_F$upper.CL)

#
# Application

# Effect sizes
at <- list("Fertiliser"="urea+", "Method"="Micromet")
m1_est_A <- emmeans(m1_lmer_p1, ~ Application, at=at)
m1_est_A <- data.frame(m1_est_A)

m1_est_A$emmean <- as.numeric(as.character(m1_est_A$emmean))
m1_est_A$emmean <- inv_signed_cube_root(m1_est_A$emmean)
m1_est_A$lower.CL <- inv_signed_cube_root(m1_est_A$lower.CL)
m1_est_A$upper.CL <- inv_signed_cube_root(m1_est_A$upper.CL)

#
# Method

# Effect sizes
at <- list("Fertiliser"="urea+", "Application"="Broadcast")
m1_est_M <- emmeans(m1_lmer_p1, ~ Method, at=at)
m1_est_M <- data.frame(m1_est_M)

m1_est_M$emmean <- as.numeric(as.character(m1_est_M$emmean))
m1_est_M$emmean <- inv_signed_cube_root(m1_est_M$emmean)
m1_est_M$lower.CL <- inv_signed_cube_root(m1_est_M$lower.CL)
m1_est_M$upper.CL <- inv_signed_cube_root(m1_est_M$upper.CL)

#

## 
# Collating a dataframe with all effect sizes

# Vectors for emmeans-objects and fectors
Model_vec <- c("m1_est_F", "m1_est_A", "m1_est_M")
Factor_vec <- c("Fertiliser","Application", "Method")

# Creating an empty dataframe
m1_estimates <- data.frame(Factor=factor(), Level=factor(), Effect=numeric(), SE=numeric(), df=numeric(), lower.CL=numeric(), upper.CL=numeric())

# Loop for extracting all estimates
for(i in c(1:3)){ # Cycling through models
  m_est <- data.frame(get(Model_vec[i]))
  m1_estimates <- rbind(m1_estimates, data.frame(Factor=Factor_vec[i],
                                                 Level=m_est[,1],
                                                 Effect=m_est[,2],
                                                 SE=m_est[,3],
                                                 df=m_est[,4],
                                                 lower.CL=m_est[,5],
                                                 upper.CL=m_est[,6]))
}
m1_estimates

#write.csv(m1_estimates, "C:/Users/au721810/Projects/2022, NH3 loss, Hutchings/Analysis/2022_NH3loss_Hutchings/output/Model_estimates/2022_NH3loss_Model1_Effects.csv")

```

The same effect sizes (and their 95% confidence interval) are shown below:

```{r echo=FALSE, include=TRUE}
# Full figure 
ggplot(m1_estimates, aes(y=Level, x=Effect, xmin=lower.CL, xmax=upper.CL, colour=Factor)) +
  theme_classic(base_size = 11) +
  facet_grid(Factor ~ ., scales = "free_y", space = "free_y") +
  geom_point(size=3) + 
  geom_errorbarh(height=.1)  +
  geom_vline(xintercept=0, colour="black", linetype="dashed") +
  scale_colour_manual(values=c("firebrick3", "yellow4", "mediumblue", "palegreen4")) +
  #scale_color_brewer(palette = "Dark2") +
  labs(x="Effect Size (NH3 loss (%))", y="") +
  theme(legend.position = "none")
```

\normalsize

Model A_Base explains `r round(r.squaredGLMM(m1_lmer_p1)[2]*100, 2)` % of the variation in the data (conditional R2), with the fixed effects explaining `r round(r.squaredGLMM(m1_lmer_p1)[1]*100, 2)` % (marginal R2). The R2-values for linear mixed effects model were calculated according to Nagakawa & Schielzeth (2013).

```{r include=FALSE, echo=FALSE}

##
# Model explanatory power
#

#
# R-squared
r.squaredGLMM(m1_lmer_p1)
RMSE_cube <- merTools::RMSE.merMod(m1_lmer_p1)
AIC_cube <- AIC(m1_lmer_p1)

#
#
# 

# Save fitted values for future models
NH3predict$M1_fit <- fitted(m1_lmer_p1)
NH3predict$M1_res <- resid(m1_lmer_p1) # For R2-calculations
```

\newpage

## Model A_Application rate

In model A_Application rate, as in all models following, we applied a linear model using fitted values from Model A_Base as an offset variable, to account for the variation explained by the factors already estimated in Model 1. As in Model A_Base, we transformed the response variable (NH3-loss %) with a squareroot-transformation with a constant of 4. The model was also weighted the analysis with the same weights as Model A_Base to account for heteroscedasticity. This same structure was used in all following models.

The sole explanatory variable in model A_Application rate was application rate (kg/ha).

```{r echo=FALSE, include=FALSE}

#########################################################################################################################################################
## MODEL 2

# 
# 
NH3predict_m2 <- filter(NH3predict, is.na(Application.rate)==F)

#
#
```

```{r include=TRUE, echo=FALSE}
hist(NH3predict_m2$Application.rate, breaks=50)

```

The model as it was applied can be seen below:

```{r include=TRUE, echo=TRUE}
# Model A_Application rate
m2_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Application.rate + offset(M1_fit),
            weights=Weight,
            data=NH3predict_m2)
```

The model assumptions can be seen below:

```{r include=TRUE, echo=FALSE}
# Model assumptions
par(mfrow=c(1,3))
plot(fitted(m2_lm), NH3predict_m2$Weight*resid(m2_lm))
abline(h=0)
qqnorm(resid(m2_lm))
qqline(resid(m2_lm))
hist(resid(m2_lm))
par(mfrow=c(1,1))
```

Application rate did not significantly explain variation in NH3-emissions (F(`r anova(m2_lm)$Df[1]`, `r anova(m2_lm)$Df[2]`) =`r round(anova(m2_lm)$F[1], 2)`, p `r ifelse(round(anova(m2_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(m2_lm)$P[1], 3)))`). There was a statistically non-significant positive trend between Application rate and NH3-emissions ($\beta$+/-SE = `r round(coef(m2_lm)[[2]],5)` +/- `r round(summary(m2_lm)$coef[2,2],5)`).

Full model output can be seen below:

```{r, include=TRUE, echo=FALSE}
# Results
summary(m2_lm)

# Save fitted values for plotting
NH3predict_m2$M2_fits <- fitted(m2_lm)

```

```{r include=FALSE, echo=FALSE}

# Extracting correct R2-measure (including the effect of offset)

m2_lm_res <- lm(M1_res ~ Application.rate,
            weights=Weight,
            data=NH3predict_m2)

# Adjusted R2
round(summary(m2_lm_res)$adj.r.squared*100, 2)

```

Application rate explains `r round(summary(m2_lm_res)$adj.r.squared*100, 2)` % of residual variation in NH3-emissions (adjusted R-squared). (Note: Adjusted R-squared is manually adjusted to incorporate the offset-variable, and thus not consistent with the adjusted R-squared from the above model output.)

Figure below is a linear approximation of the effect of Application rate on NH3-emissions (linear slope back-transformed from the square-root scale).

```{r include=TRUE, echo=FALSE, message=FALSE}
#
#
#

# Model figure

ggplot(NH3predict_m2, aes(y=NH3loss, x=Application.rate)) + 
  theme_classic(base_size = 11) +
  geom_point(position=position_jitter(1), colour="gray75") +
  geom_smooth(aes(y=(M2_fits^2)-4), method='lm', se=T, colour="black") +
  labs(x="Application rate (kg/ha)", y = "NH3 loss (%)") 

#
#
# 
```

\newpage

## Model A_Soil pH

```{r include=FALSE, echo=FALSE}

#########################################################################################################################################################
## MODEL AA_Soil pH

#
#
# Remove NA's from the dataset - n=1894
NH3predict_m3 <- NH3predict %>% drop_na(SoilpH)
# NH3predict_m3<-NH3predict_m3 %>%  mutate(pHgroup = case_when(SoilpH < 6.0 ~ 'low',
#                            SoilpH >=6.0 & SoilpH<7.0 ~ 'norm',
#                            SoilpH >=7.0 & SoilpH <8.0 ~ 'high',
#                            SoilpH >8.0 ~ 'very_high'))
# NH3predict_m3$pHgroup<-factor(NH3predict_m3$pHgroup)

```

Soil pH was binarised into categories, where soils with pH\>7.0 were categorised as chalky, and ph\<=7.0 as normal. The number of observations in each group and their frequency distribution can be seen below.

```{r include=TRUE, echo=FALSE}

ddply(NH3predict_m3, c("pH_cat"), summarise, N=length(NH3loss))
ggplot(NH3predict_m3, aes(x=SoilpH, fill=pH_cat)) + 
  theme_classic(base_size=11) + 
  geom_histogram(binwidth=0.1, alpha=.5, position="identity")

```

Model A_Soil pH followed the structure of Model A_Application. We added Soil pH, Fertiliser type, and the interaction between the two as fixed factors. In this model, Fertiliser types were used to identify differences between pH-categories, not to analyse the effect of Fertiliser on its own - the offset of Model A_Base removes the variation explained by Fertilisers otherwise.

```{r include=TRUE, echo=TRUE}

# Model
m3_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Fertiliser * pH_cat + offset(M1_fit), 
            weights=Weight,
            data=NH3predict_m3)

```

The heteroscedasticity and normality of errors were as below:

```{r include=TRUE, echo=FALSE}

# Model assumptions
par(mfrow=c(1,3))
plot(fitted(m3_lm), NH3predict_m3$Weight*resid(m3_lm))
abline(h=0)
qqnorm(resid(m3_lm))
qqline(resid(m3_lm))
hist(resid(m3_lm))
par(mfrow=c(1,1))

```

The effect of soil pH on NH3-emissions was dependent on the Fertiliser types (F(`r anova(m3_lm)$Df[3]`, `r anova(m3_lm)$Df[4]`) =`r round(anova(m3_lm)$F[3], 2)`, p `r ifelse(round(anova(m3_lm)$P[3], 3)=="0", "<0.001", paste("=",round(anova(m3_lm)$P[3], 3)))`). Soil pH also had an independent effect on NH3 emissions (F(`r anova(m3_lm)$Df[2]`, `r anova(m3_lm)$Df[4]`) =`r round(anova(m3_lm)$F[2], 2)`, p `r ifelse(round(anova(m3_lm)$P[2], 3)=="0", "<0.001", paste("=",round(anova(m3_lm)$P[2], 3)))`).

Full anova-table can be seen below:

```{r include=TRUE, echo=FALSE}

anova(m3_lm)

```

Full model output can be seen below:

```{r include=TRUE, echo=FALSE}
# Results
summary(m3_lm)

# Save fitted values for plotting
NH3predict_m3$pH_fits <- fitted(m3_lm)


```

```{r include=TRUE, echo=FALSE}

# Extracting correct R2-measure (including the effect of offset)

m3_lm_res <- lm(M1_res ~ -1 + Fertiliser:pH_cat, 
                weights=Weight,
            data=NH3predict_m3)


```

The model explained `r round(summary(m3_lm_res)$adj.r.squared*100, 2)` % of the residual variation in NH3-loss.

### Combining A_Base and A_Soil pH

For acquiring estimates for Soil pH in different fertiliser types at the same time, and same scale as the base-model, we added Soil pH and the interaction between pH and FertiliserType into the initial model. Please note this model has less data than Model A_Base, and as it includes new parameters, the estimates change slightly.

```{r include=FALSE, echo=TRUE}

m3_lmer <- lmer(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Fertiliser + Application + Method + pH_cat + Fertiliser:pH_cat + (1|Ref),
                   weights=Weight,
                   control=lmerControl(optimizer="bobyqa"),
                   data=NH3predict)
summary(m3_lmer)
```

The effect sizes (marginal means) for the interaction can be seen below. The effect sizes are given at the back-transformed scale (NH3-loss %), parametrised to Fertiliser=urea+, Application=Broadcast, Method=Micromet, pH=normal.

```{r include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#
##### 
# Estimates for each main effect


#
# Fertiliser

# Effect sizes
at <- list("Application"="Broadcast", "Method"="Micromet", "pH_cat"="normal")
m3_est_F <- emmeans(m3_lmer, ~ Fertiliser, at=at)
m3_est_F  <- data.frame(m3_est_F)

m3_est_F$emmean <- as.numeric(as.character(m3_est_F$emmean))
m3_est_F$emmean <- inv_signed_cube_root(m3_est_F$emmean)
m3_est_F$lower.CL <- inv_signed_cube_root(m3_est_F$lower.CL)
m3_est_F$upper.CL <- inv_signed_cube_root(m3_est_F$upper.CL)

#
# Application

# Effect sizes
at <- list("Fertiliser"="urea+", "Method"="Micromet", "pH_cat"="normal")
m3_est_A <- emmeans(m3_lmer, ~ Application, at=at)
m3_est_A <- data.frame(m3_est_A)

m3_est_A$emmean <- as.numeric(as.character(m3_est_A$emmean))
m3_est_A$emmean <- inv_signed_cube_root(m3_est_A$emmean)
m3_est_A$lower.CL <- inv_signed_cube_root(m3_est_A$lower.CL)
m3_est_A$upper.CL <- inv_signed_cube_root(m3_est_A$upper.CL)

#
# Method

# Effect sizes
at <- list("Fertiliser"="urea+", "Application"="Broadcast", "pH_cat"="normal")
m3_est_M <- emmeans(m3_lmer, ~ Method, at=at)
m3_est_M <- data.frame(m3_est_M)

m3_est_M$emmean <- as.numeric(as.character(m3_est_M$emmean))
m3_est_M$emmean <- inv_signed_cube_root(m3_est_M$emmean)
m3_est_M$lower.CL <- inv_signed_cube_root(m3_est_M$lower.CL)
m3_est_M$upper.CL <- inv_signed_cube_root(m3_est_M$upper.CL)

#
# Soil pH

# Effect size
at <- list("Fertiliser"="urea+", "Application"="Broadcast", "Method"="Micromet")
m3_est_pH <- emmeans(m3_lmer, ~ pH_cat, at=at)
m3_est_pH <- data.frame(m3_est_pH)

m3_est_pH$emmean <- as.numeric(as.character(m3_est_pH$emmean))
m3_est_pH$emmean <- inv_signed_cube_root(m3_est_pH$emmean)
m3_est_pH$lower.CL <- inv_signed_cube_root(m3_est_pH$lower.CL)
m3_est_pH$upper.CL <- inv_signed_cube_root(m3_est_pH$upper.CL)

#
# Interaction

# Effect size
at <- list("Application"="Broadcast", "Method"="Micromet")
m3_est_pHFe <- emmeans(m3_lmer, pairwise ~ pH_cat | Fertiliser, at=at)
m3_est_pHFe <- data.frame(m3_est_pHFe)

m3_est_pHFe$emmean <- as.numeric(as.character(m3_est_pHFe$emmean))
m3_est_pHFe$emmean <- inv_signed_cube_root(m3_est_pHFe$emmean)
m3_est_pHFe$lower.CL <- inv_signed_cube_root(m3_est_pHFe$lower.CL)
m3_est_pHFe$upper.CL <- inv_signed_cube_root(m3_est_pHFe$upper.CL)


#

## 
# Collating a dataframe with all effect sizes

# Vectors for emmeans-objects and fectors
Model_vec <- c("m3_est_F", "m3_est_A", "m3_est_M", "m3_est_pH")
Factor_vec <- c("Fertiliser", "Application", "Method", "Soil pH")

# Creating an empty dataframe
m3_estimates <- data.frame(Factor=factor(), Level=factor(), Effect=numeric(), SE=numeric(), df=numeric(), lower.CL=numeric(), upper.CL=numeric())

# Loop for extracting all estimates
for(i in c(1:4)){ # Cycling through models
  m_est <- data.frame(get(Model_vec[i]))
  m3_estimates <- rbind(m3_estimates, data.frame(Factor=Factor_vec[i],
                                                 Level=m_est[,1],
                                                 Effect=m_est[,3],
                                                 SE=m_est[,4],
                                                 df=m_est[,5],
                                                 lower.CL=m_est[,5],
                                                 upper.CL=m_est[,6]))
}


m3_est_pHFe$Level <- paste(m3_est_pHFe$Fertiliser, ":", m3_est_pHFe$pH_cat)


# Add the interaction
m3_estimates <- rbind(m3_estimates,
                      data.frame(Factor="Fertiliser:Soil pH",
                                 Level=filter(m3_est_pHFe, 
                                              contrast==".")$Level,
                                 Effect=filter(m3_est_pHFe,
                                          contrast==".")$emmean,
                                 SE=filter(m3_est_pHFe,
                                           contrast==".")$SE,
                                 df=filter(m3_est_pHFe,
                                           contrast==".")$df,
                                 lower.CL=filter(m3_est_pHFe,
                                          contrast==".")$lower.CL,
                                 upper.CL=filter(m3_est_pHFe,
                                          contrast==".")$upper.CL))

m3_estimates

#write.csv(m3_estimates, "C:/Users/au721810/Projects/2022, NH3 loss, Hutchings/Analysis/2022_NH3loss_Hutchings/output/Model_estimates/2022_NH3loss_Model1_EffectsM13.csv")

```

The same effect sizes (and their 95% confidence interval) are shown below:

```{r include=TRUE, echo=FALSE}

# Figure
#ggplot(m3_estimates, aes(y=Level, x=Effect, xmin=lower.CL, xmax=upper.CL,

ggplot(m3_estimates, aes(y=Level, x=Effect, xmin=lower.CL, xmax=upper.CL, colour=Factor))+
#xlim(0, 30) +
theme(strip.text.y = element_text(size = 6)) +
  facet_grid(Factor ~ ., scales = "free_y", space = "free_y") +
  geom_point(size=3) + 
  geom_errorbarh(height=.1)  +
  geom_vline(xintercept=0, colour="black", linetype="dashed") +
  scale_color_brewer(palette = "Dark2") +
  labs(x="Effect Size (NH3 loss (%))", y="") +
  theme(legend.position = "none")

 
```

\newpage

## Model A_Soil clay content

In Model A_Soil clay content, we explored the effect of Soil clay content (%) on NH3-loss (%).

```{r include=FALSE, echo=FALSE}

#########################################################################################################################################################
## MODEL 4

#
#
# Remove NA's from the dataset - n=1894
NH3predict_m4 <- NH3predict %>% drop_na(Clay)
#NH3ClayOut<-NH3predict_m4[c('Clay', 'NH3loss')]
#write.csv(NH3ClayOut,"Model4_data_all.csv")
NH3predict_m4b <- NH3predict %>% drop_na(SoilCEC)

```

```{r include=TRUE, echo=FALSE}

hist(NH3predict_m4$Clay, breaks=50)

```

We used Soil clay content (%) as a continuous covariate and the sole explanatory variable in the model.

```{r include=TRUE, echo=TRUE}
# Model
m4_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Clay + offset(M1_fit), 
            weights=Weight, 
            data=NH3predict_m4)

```

The model assumptions (heteroscedasticity and normality of errors):

```{r include=TRUE, echo=FALSE}

# Model assumptions
par(mfrow=c(1,3))
plot(fitted(m4_lm), NH3predict_m4$Weight*resid(m4_lm))
abline(h=0)
qqnorm(resid(m4_lm))
qqline(resid(m4_lm))
hist(resid(m4_lm))
par(mfrow=c(1,1))

```

Soil clay content (%) affected NH3-loss (%) negatively ($\beta$+/-SE = `r round(coef(m4_lm)[[2]],5)` +/- `r round(summary(m4_lm)$coef[2,2],5)`, F(`r anova(m4_lm)$Df[1]`, `r anova(m4_lm)$Df[2]`) =`r round(anova(m4_lm)$F[1], 2)`, p `r ifelse(round(anova(m4_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(m4_lm)$P[1], 3)))`).

Full model output can be seen below:

```{r include=TRUE, echo=FALSE}
# Results
summary(m4_lm)

# Save fitted values for plotting
NH3predict_m4$M4_fits <- fitted(m4_lm)

```

```{r include=FALSE, echo=FALSE}

# Extracting correct R2-measure (including the effect of offset)

m4_lm_res <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Clay, 
            weights=Weight, 
            data=NH3predict_m4)

round(summary(m4_lm_res)$adj.r.squared*100, 2)
```

The model explained `r round(summary(m4_lm_res)$adj.r.squared*100, 2)` % of the residual variation in NH3-loss.

Figure below is a linear approximation of the effect of Soil Clay content (%) on NH3-emissions (linear slope back-transformed from the square-root scale).

```{r include=TRUE, echo=FALSE, message=FALSE}
# Model figure
ggplot(NH3predict_m4, aes(y=NH3loss, x=Clay)) + 
  theme_classic(base_size = 11) +
  geom_point(colour="gray75") +
  geom_smooth(aes(y=(M4_fits^2)-4), method='lm', se=T, colour="black") +
  labs(x="Soil clay content (%)", y = "NH3 loss (%)") 

```

Test to see if a exponential function provides a better fit for the Clay data. The fit and parameters c and d were fitted a non-linear least-squares model.

```{r include=TRUE, echo=FALSE}

library(minpack.lm)
MR_nls_clay <- nlsLM(NH3loss ~ c * exp(d*(Clay)), 
                   data=NH3predict_m4, 
                   start=list(c=0, d=0.05),
                   control = list(maxiter = 100, scale = "levenberg"),
                   algorithm="ptol")
summary(MR_nls_clay)
```

The fitted relationship is as below:

```{r include=TRUE, echo=FALSE}
ggplot(NH3predict_m4, aes(x=Clay, y=NH3loss))+
  theme_bw() +
  geom_point(colour="gray75") +
  geom_line(stat = "smooth", 
            method = "nlsLM",
            formula =  y ~ c * exp(d*(x)),
            method.args=list(start=list(c=0, d=0.05),
                             control = list(maxiter = 100, scale = "levenberg")),                 
            colour="black",
            linetype="solid",
            size=1,
            se = FALSE) +
  labs(x="Clay (%))", y="sqrt(NH3 loss (%))")


```

```{r include=TRUE, echo=TRUE}

# Transforming the Clay values to exp-scale
NH3predict_m4$Clay.exp <- coef(MR_nls_clay)[[1]]*exp(coef(MR_nls_clay)[[2]]*NH3predict_m4$Clay)

```

```{r include=TRUE, echo=FALSE}
#
# Model A_Soil clay content_b
mTR_lm_expClay <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~Clay.exp + offset(M1_fit), 
             weights=Weight, 
             data=NH3predict_m4)

# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(mTR_lm_expClay))
qqline(resid(mTR_lm_expClay))
plot(fitted(mTR_lm_expClay), NH3predict_m4$Weight*resid(mTR_lm_expClay))
abline(h=0)
hist(resid(mTR_lm_expClay))
par(mfrow=c(1,1))

# Results
summary(mTR_lm_expClay)
```

\newpage

## Model A_Soil organic carbon

Soil organic carbon (g/kg) was modeled as a continuous variable. Most soils in the data were mineral (with a cutoff of 120 g/kg SOC), with only 3 organic soils in the dataset.

```{r include=FALSE, echo=FALSE}
#
#
# Remove NA's from the dataset - n=1894
NH3predict_m5 <- NH3predict %>% drop_na(SOC)


```

```{r include=TRUE, echo=FALSE}

hist(NH3predict_m5$SOC, breaks=59)

```

We used Soil organic carbon (g/kg) as a sole continuous covariate in the model.

```{r include=TRUE, echo=TRUE}

# Model
m5_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ SOC + offset(M1_fit), 
            weights=Weight,
            data=NH3predict_m5)

```

The model assumptions (heteroscedasticity and normality of errors):

```{r include=TRUE, echo=FALSE}

# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(m5_lm))
qqline(resid(m5_lm))
plot(fitted(m5_lm), NH3predict_m5$Weight*resid(m5_lm))
abline(h=0)
hist(resid(m5_lm))
par(mfrow=c(1,1))

```

There was no effect of Soil organic carbon (g/kg) on NH3 loss (F(`r anova(m5_lm)$Df[1]`, `r anova(m5_lm)$Df[2]`) =`r round(anova(m5_lm)$F[1], 2)`, p `r ifelse(round(anova(m5_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(m5_lm)$P[1], 3)))`).

Full model output can be seen below:

```{r include=TRUE, echo=FALSE, message=FALSE}
# Results
summary(m5_lm)

# Save fitted values for plotting
NH3predict_m5$M5_fits <- fitted(m5_lm)

```

```{r include=FALSE, echo=FALSE}

# Extracting correct R2-measure (including the effect of offset)

m5_lm_res <- lm(M1_res ~ SOC, 
            weights=Weight,
            data=NH3predict_m5)

round(summary(m5_lm_res)$adj.r.squared*100, 2)
anova(m5_lm_res)

round(summary(m5_lm_res)$r.squared*100, 2)
1-(1544.54/(1544.54+0.03))

```

Adjusted R2-value of the model was `r round(summary(m5_lm_res)$adj.r.squared, 3)`, and multiple R2 was `r round(summary(m5_lm_res)$r.squared,6)`. The negative adjusted R2-value indicates that the amount of variation accounted for by Soil organic carbon is negligible. In practice, soil organic carbon explains no variation in the data.

Figure below is a linear approximation of the effect of Soil organic carbon (g/kg) on NH3-emissions (linear slope back-transformed from the square-root scale).

```{r include=TRUE, echo=FALSE, message=FALSE}
#
#
#

# Model figure
ggplot(NH3predict_m5, aes(y=NH3loss, x=SOC)) + 
  theme_classic(base_size = 11) +
  geom_point(colour="gray75") +
  geom_smooth(aes(y=(M5_fits^2)-4), method='lm', se=T, colour="black") +
  labs(x="Soil organic carbon (g/kg)", y = "NH3 loss (%)") 

```

\newpage

## Model A_Cover

Crop cover was split into three categories: no cover ("none"), arable crops (including alfalfa, cereal, maize, rape and winter cereal), and grass. The number of observations in each group, and their NH3-loss (%) distributions can be seen below:

```{r include=FALSE, echo=FALSE}
#
#
# Remove NA's from the dataset - n=1894
NH3predict_m6 <- NH3predict %>% drop_na(Cover)
NH3predict_m6 <- filter(NH3predict_m6, Crop!="rice" & Cover!="other")

# Removing empty factor levels
NH3predict_m6$Cover <- factor(NH3predict_m6$Cover)
# Reordering the factor levels
NH3predict_m6$Cover <- relevel(NH3predict_m6$Cover, "None")


```

```{r include=TRUE, echo=FALSE}

ddply(NH3predict_m6, c("Cover"), summarise, N=length(NH3loss))

ggplot(NH3predict_m6, aes(x=NH3loss, fill=Cover)) + 
  theme_classic(base_size=11) +
  facet_wrap(~Cover) + 
  geom_histogram(binwidth=5, alpha=.5, position="identity")

```

We added Cover as a fixed factor and a sole explanatory variable in the model.

```{r include=TRUE, echo=TRUE}
# Model
m6_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Cover + offset(M1_fit), 
            weights=Weight,
            data=NH3predict_m6)

```

The model assumptions (heteroscedasticity and normality of errors):

```{r include=TRUE, echo=FALSE}
# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(m6_lm))
qqline(resid(m6_lm))
plot(fitted(m6_lm), NH3predict_m6$Weight*resid(m6_lm))
abline(h=0)
hist(resid(m6_lm))
par(mfrow=c(1,1))

```

The cover categories (none, arable, grass) did not significantly differ from one another in NH3-emissions (F(`r anova(m6_lm)$Df[1]`, `r anova(m6_lm)$Df[2]`) =`r round(anova(m6_lm)$F[1], 2)`, p `r ifelse(round(anova(m6_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(m6_lm)$P[1], 3)))`).

Full model output can be seen below:

```{r include=TRUE, echo=FALSE}
# Results
summary(m6_lm)

# Save fitted values for plotting
NH3predict_m6$M6_fits <- fitted(m6_lm)

```

```{r include=FALSE, echo=FALSE}

# Extracting correct R2-measure (including the effect of offset)

m6_lm_res <- lm(M1_res ~ Cover, 
            weights=Weight,
            data=NH3predict_m6)

round(summary(m6_lm_res)$adj.r.squared*100, 2)

```

Adjusted R2-value of the model was `r round(summary(m6_lm_res)$adj.r.squared, 5)`, and multiple R2 was `r round(summary(m6_lm_res)$r.squared,5)`. The negative adjusted R2-value indicates that the amount of variation accounted for by Cover crops is negligible. In practice, cover crops explain no variation in the data.

The effect sizes (marginal means) of % NH3-loss for each cover category can be seen below.

```{r include=TRUE, echo=FALSE}

# Effect sizes
m6_est <- emmeans(m6_lm, ~ Cover)
m6_est <- data.frame(m6_est)

m6_est
```

```{r include=TRUE, echo=FALSE}
# Model figure
ggplot(m6_est, aes(y=Cover, x=emmean, xmin=lower.CL, xmax=upper.CL)) +
  theme_classic(base_size = 11) +
  geom_point(size=3) + 
  geom_errorbarh(height=.1)  +
  geom_vline(xintercept=0, colour="black", linetype="dashed") +
  labs(x="Effect Size (NH3 loss (%))", y="") +
  theme(legend.position = "none")

```

\newpage

## Model A_Air temperature and rainfall

AirTemperature over the duration of the experiment (Celcius) and mean rainfall over the experimental period (mm) were analysed together, to explore interactions between the two variables. The frequency distributions of the two variables can be seen below.

```{r include=FALSE, echo=FALSE}
#########################################################################################################################################################
## MODEL A7-A8

#
#
# Remove NA's from the dataset
NH3predict_m7 <- NH3predict %>% drop_na(c(AirTemperature, Rain.mean))

NH3predict_m7 <- filter(NH3predict_m7, Place=="Outdoor")

cor(NH3predict_m7$AirTemperature,NH3predict_m7$Rainfall)
cor.test(NH3predict_m7$AirTemperature,NH3predict_m7$Rainfall)
```

There are `r length(NH3predict_m7$NH3loss)` observations in the analysis.

```{r include=TRUE, echo=FALSE}

hist(NH3predict_m7$AirTemperature, breaks=50)
hist(NH3predict_m7$Rain.mean, breaks=50)

```

```{r include=TRUE, echo=TRUE}
# Model
mTR_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ AirTemperature * Rain.mean + offset(M1_fit), 
             weights=Weight, 
             data=NH3predict_m7)

```

The model assumptions (heteroscedasticity and normality of errors):

```{r include=TRUE, echo=FALSE}

# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(mTR_lm))
qqline(resid(mTR_lm))
plot(fitted(mTR_lm), NH3predict_m7$Weight*resid(mTR_lm))
abline(h=0)
hist(resid(mTR_lm))
par(mfrow=c(1,1))


```

There was a marginally non-significant effect of interaction (F(`r anova(mTR_lm)$Df[3]`, `r anova(mTR_lm)$Df[4]`) =`r round(anova(mTR_lm)$F[3], 2)`, p `r ifelse(round(anova(mTR_lm)$P[3], 3)=="0", "<0.001", paste("=",round(anova(mTR_lm)$P[3], 3)))`), and a marginally non-significant main effect of mean rainfall (F(`r anova(mTR_lm)$Df[2]`, `r anova(mTR_lm)$Df[4]`) =`r round(anova(mTR_lm)$F[2], 2)`, p `r ifelse(round(anova(mTR_lm)$P[2], 3)=="0", "<0.001", paste("=",round(anova(mTR_lm)$P[2], 3)))`). There was no main effect of air temperature (F(`r anova(mTR_lm)$Df[1]`, `r anova(mTR_lm)$Df[4]`) =`r round(anova(mTR_lm)$F[1], 2)`, p `r ifelse(round(anova(mTR_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(mTR_lm)$P[1], 3)))`).

Full model output can be seen below:

```{r include=TRUE, echo=FALSE}
# Results
summary(mTR_lm)

# Save fitted values for plotting
NH3predict_m7$M7_fits <- fitted(mTR_lm)


```

```{r include=FALSE, echo=FALSE}

# Extracting correct R2-measure (including the effect of offset)

mTR_lm_res <- lm(M1_res ~ AirTemperature * Rain.mean, 
             weights=Weight, 
             data=NH3predict_m7)

round(summary(mTR_lm_res)$adj.r.squared*100, 2)

```

Air temperature and rainfall, and the interaction between the two explain `r round(summary(mTR_lm_res)$adj.r.squared*100, 2)` % of the variation in the data.

The marginally non-significant interaction between rainfall and temperature can be interpreted as the slope of air temperature increasing in steepness with increasing rainfall. This means that the effect of temperature is dependent on rainfall, and temperature only has an effect when rain is present.

```{r include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

# Rainfall modifying Temperature
interact_plot(mTR_lm, pred=AirTemperature, modx=Rain.mean,
              partial.residuals = TRUE, 
              #linearity.check = TRUE,
              modx.values = c(0, 5,10,15,20,25))

```

Similarly, the interaction can be interpreted from the perspective of rain: The association between rain and NH3 emissions changes from negative to slightly positive with increasing temperature.

```{r include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

# Temperature modifying Rainfall
interact_plot(mTR_lm, pred=Rain.mean, modx=AirTemperature, 
              partial.residuals = TRUE, 
              modx.values = c(0, 5, 10, 15, 20, 25, 30, 35))


```

Simple slope analysis on the interaction can be seen below. The slope of temperature is statistically significant within range noted below. Note that the results below rely on statistical significance of the interaction in the model above, which is not fulfilled.

```{r include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

# Simple slopes analysis
mTR_slopes <- sim_slopes(mTR_lm, pred = AirTemperature, modx = Rain.mean, 
                         modx.values = c(0, 5,10,15,20,25))
#mTR_slopes
plot(mTR_slopes)
johnson_neyman(mTR_lm, pred = AirTemperature, modx = Rain.mean, alpha = .05)

```

### A_Air temperature

```{r include=FALSE, echo=TRUE}
NH3predict_m7_2 <- NH3predict %>% drop_na(c(AirTemperature))
##NH3predict_m7_2 <- filter(NH3predict_m7_2, Place=="Outdoor")


m7_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ AirTemperature + offset(M1_fit), 
             weights=Weight, 
             data=NH3predict_m7_2)

m7_lm_res <- lm(M1_res ~ AirTemperature, 
             weights=Weight, 
             data=NH3predict_m7_2)

# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(m7_lm))
qqline(resid(m7_lm))
plot(fitted(m7_lm), NH3predict_m7_2$Weight*resid(m7_lm))
abline(h=0)
hist(resid(m7_lm))
par(mfrow=c(1,1))

NH3predict_m7_2$M7_fits <- fitted(m7_lm)


```

Considering air temperature separately, there is a statistically significant effect of air temperature (F(`r anova(m7_lm)$Df[1]`, `r anova(m7_lm)$Df[2]`) =`r round(anova(m7_lm)$F[1], 2)`, p `r ifelse(round(anova(m7_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(m7_lm)$P[1], 3)))`). Adjusted R2-value of the model was `r round(summary(m7_lm_res)$adj.r.squared, 5)`, and multiple R2 was `r round(summary(m7_lm_res)$r.squared,5)`. This means air temperature explains `r round(summary(m7_lm_res)$adj.r.squared, 5)*100` % of the variation.

There are `r length(NH3predict_m7_2$NH3loss)` observations in the analysis.

The full model output with the slopes and their standard error can be seen below:

```{r include=TRUE, echo=FALSE}

summary(m7_lm)

```

### A_Air temperature over first 7 days

```{r include=FALSE, echo=TRUE}
NH3predict_mA7 <- NH3predict %>% drop_na(c(Temp.7d))
NH3predict_mA7$Temp.7d <-  as.numeric(as.character(NH3predict_mA7$Temp.7d))
NH3predict_mA7 <- NH3predict_mA7 %>% drop_na(c(Temp.7d))
#NH3predict_mA7 <- filter(NH3predict_mA7, Place=="Outdoor")

NH3predict_mA7$Temp.7d <- as.numeric(as.character(NH3predict_mA7$Temp.7d))
mA7_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Temp.7d + offset(M1_fit), 
             weights=Weight, 
             data=NH3predict_mA7)

mA7_lm_res <- lm(M1_res ~ Temp.7d, 
             weights=Weight, 
             data=NH3predict_mA7)

# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(mA7_lm))
qqline(resid(mA7_lm))
plot(fitted(mA7_lm), NH3predict_mA7$Weight*resid(mA7_lm))
abline(h=0)
hist(resid(mA7_lm))
par(mfrow=c(1,1))

NH3predict_mA7$MA7_fits <- fitted(mA7_lm)

```

There is no statistically significant effect of air temperature over the first 7 days of the experiment (F(`r anova(mA7_lm)$Df[1]`, `r anova(mA7_lm)$Df[2]`) =`r round(anova(mA7_lm)$F[1], 2)`, p `r ifelse(round(anova(mA7_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(mA7_lm)$P[1], 3)))`). Adjusted R2-value of the model was `r round(summary(mA7_lm_res)$adj.r.squared, 5)`, and multiple R2 was `r round(summary(mA7_lm_res)$r.squared,5)`. This means air temperature over the first 7 days explains `r round(summary(mA7_lm_res)$adj.r.squared, 5)*100` % of the variation.

There are `r length(NH3predict_mA7$NH3loss)` observations in the analysis.

The full model output with the slopes and their standard error can be seen below:

```{r include=TRUE, echo=FALSE}

summary(mA7_lm)

```

### A_Mean daily rainfall

```{r include=FALSE, echo=TRUE}
NH3predict_m8 <- NH3predict %>% drop_na(c(Rain.mean))
NH3predict_m8 <- filter(NH3predict_m8, Place=="Outdoor")

m8_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Rain.mean + offset(M1_fit), 
             weights=Weight, 
             data=NH3predict_m8)

m8_lm_res <- lm(M1_res ~ Rain.mean, 
             weights=Weight, 
             data=NH3predict_m8)

# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(m8_lm))
qqline(resid(m8_lm))
plot(fitted(m8_lm), NH3predict_m8$Weight*resid(m8_lm))
abline(h=0)
hist(resid(m8_lm))
par(mfrow=c(1,1))

NH3predict_m8$M8_fits <- fitted(m8_lm)


```

Considering mean rainfall separately, there is no statistically significant effect of rain (F(`r anova(m8_lm)$Df[1]`, `r anova(m8_lm)$Df[2]`) =`r round(anova(m8_lm)$F[1], 2)`, p `r ifelse(round(anova(m8_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(m8_lm)$P[1], 3)))`). Adjusted R2-value of the model was `r round(summary(m8_lm_res)$adj.r.squared, 5)`, and multiple R2 was `r round(summary(m8_lm_res)$r.squared,5)`. Thus, rainfall doesn't meaningfully explain any variation in the data.

There are `r length(NH3predict_m8$NH3loss)` observations in the analysis.

The full model output with the slopes and their standard error can be seen below:

```{r include=TRUE, echo=FALSE}

summary(m8_lm)

```

### A_Rainfall over first 72 h

```{r include=FALSE, echo=TRUE}
NH3predict_R8 <- NH3predict %>% drop_na(c(Rain.72h))
NH3predict_R8 <- filter(NH3predict_R8, Place=="Outdoor")

NH3predict_R8$Rain.72h <- as.numeric(as.character(NH3predict_R8$Rain.72h))
NH3predict_R8 <- NH3predict_R8 %>% drop_na(c(Rain.72h))

NH3predict$Rain.72h <- as.numeric(as.character(NH3predict$Rain.72h))

mR8_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Rain.72h + offset(M1_fit), 
             weights=Weight, 
             data=NH3predict_R8)

mR8_lm_res <- lm(M1_res ~ Rain.72h, 
             weights=Weight, 
             data=NH3predict_R8)

# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(mR8_lm))
qqline(resid(mR8_lm))
plot(fitted(mR8_lm), NH3predict_R8$Weight*resid(mR8_lm))
abline(h=0)
hist(resid(mR8_lm))
par(mfrow=c(1,1))

NH3predict_R8$MR8_fits <- fitted(mR8_lm)


```

There is no statistically significant effect of rainfall during the first 72h of the experiment (F(`r anova(mR8_lm)$Df[1]`, `r anova(mR8_lm)$Df[2]`) =`r round(anova(mR8_lm)$F[1], 2)`, p `r ifelse(round(anova(mR8_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(mR8_lm)$P[1], 3)))`). Adjusted R2-value of the model was `r round(summary(mR8_lm_res)$adj.r.squared, 5)`, and multiple R2 was `r round(summary(mR8_lm_res)$r.squared,5)`. Thus, rainfall during the first 72h doesn't meaningfully explain any variation in the data.

There are `r length(NH3predict_R8$NH3loss)` observations in the analysis.

The full model output with the slopes and their standard error can be seen below:

```{r include=TRUE, echo=FALSE}

summary(mR8_lm)

```

Figures below show a linear approximation of the effect of air temperature (Celcius) and mean rainfall (mm) on NH3-emissions (linear slope back-transformed from the square-root scale).

```{r include=TRUE, echo= FALSE, message=FALSE}

# Figure of Temperature and Rainfall as estimated by the model
ggplot(NH3predict_m7_2, aes(x = AirTemperature, y = NH3loss)) +
  theme_classic(base_size=11) +
  geom_point(colour="gray75") +
  stat_smooth(aes(y=(M7_fits^2)-4), method="lm", colour="black") +
  labs(x="Air temperature", y="NH3 loss (%)")

ggplot(NH3predict_m8, aes(x = Rain.mean, y = NH3loss)) +
  theme_classic(base_size=11) +
  geom_point(colour="gray75") +
  stat_smooth(aes(y=(M8_fits^2)-4), method="lm", colour="black") +
  labs(x="Rainfall (mean mm/day)", y="NH3 loss (%)")

```

### A_Rainfall as a non-linear effect

We fitted an exponential function with parameters a and b to rainfall data. The equation is as follows:

$f(Rain.mean) = a * e^{(b*Rain.mean)}$

The response variable is presented in square-root(+4) scale to make the data comparable for further analysis.

To test the fit and to estimate parameters a and b, we fitted a non-linear least-squares model.

```{r include=TRUE, echo=FALSE}

library(minpack.lm)
MR_nls <- nlsLM(NH3loss ~ a * exp(b*(Rain.mean)), 
                   data=NH3predict_m7, 
                   start=list(a=0, b=0.05),
                   control = list(maxiter = 100, scale = "levenberg"),
                   algorithm="ptol")
summary(MR_nls)
```

The fitted relationship is as below:

```{r include=TRUE, echo=FALSE}
ggplot(NH3predict_m7, aes(x=Rain.mean, y=NH3loss))+
  theme_bw() +
  geom_point(colour="gray75") +
  geom_line(stat = "smooth", 
            method = "nlsLM",
            formula =  y ~ a * exp(b*(x)),
            method.args=list(start=list(a=0, b=0.05),
                             control = list(maxiter = 100, scale = "levenberg")),                 
            colour="black",
            linetype="solid",
            size=1,
            se = FALSE) +
  labs(x="Rainfall (mean mm/day)", y="NH3 loss (%)")

```

The RMSE (root mean square error) of the model was `r sqrt(mean((NH3predict_m7$NH3loss-fitted(MR_nls))^2))`, which is on the same scale as mean NH3loss values in the data (`r mean(NH3predict_m7$NH3loss)`) indicating a poor fit of the model. (Error equal to the estimated value.)

We calculated new rain estimates on the exponential scale, with the estimated parameters taken into account, and ran the analysis with air temperature again, using the exponential Rain.mean values calculated using the coefficients (parameter estimates) from the non-linear least-squares fit above.

```{r include=TRUE, echo=TRUE}

# Transforming the Rain values to exp-scale
NH3predict_m7$Rain.exp <- coef(MR_nls)[[1]]*exp(coef(MR_nls)[[2]]*NH3predict_m7$Rain.mean)

```

```{r include=TRUE, echo=FALSE}
#
# Model
mTR_lm_exp <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ AirTemperature * Rain.exp + offset(M1_fit), 
             weights=Weight, 
             data=NH3predict_m7)

# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(mTR_lm_exp))
qqline(resid(mTR_lm_exp))
plot(fitted(mTR_lm_exp), NH3predict_m7$Weight*resid(mTR_lm_exp))
abline(h=0)
hist(resid(mTR_lm_exp))
par(mfrow=c(1,1))

# Results
summary(mTR_lm_exp)
```

```{r include=TRUE, echo=FALSE}
#
# Model
mTR_lm_exp2 <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ AirTemperature + Rain.exp + offset(M1_fit), 
             weights=Weight, 
             data=NH3predict_m7)

#anova(mTR_lm_exp, mTR_lm_exp2)


```

When the interaction is removed, we end in a model showing very similar results as the linear model above, with no statistically significant effect of rainfall (F(`r anova(mTR_lm_exp2)$Df[2]`, `r anova(mTR_lm_exp2)$Df[3]`) =`r round(anova(mTR_lm_exp2)$F[2], 2)`, p `r ifelse(round(anova(mTR_lm_exp2)$P[2], 3)=="0", "<0.001", paste("=",round(anova(mTR_lm_exp)$P[2], 3)))`), nor air temperature (F(`r anova(mTR_lm_exp2)$Df[1]`, `r anova(mTR_lm_exp2)$Df[3]`) =`r round(anova(mTR_lm_exp2)$F[1], 2)`, p `r ifelse(round(anova(mTR_lm_exp2)$P[1], 3)=="0", "<0.001", paste("=",round(anova(mTR_lm_exp)$P[1], 3)))`), but a statistically significant slope for air temperature (see below).

```{r include=TRUE, echo=FALSE}

# Results
summary(mTR_lm_exp2)


```

```{r include=TRUE, echo=FALSE}

# Transforming the Rain values to exp-scale
NH3predict_m8$Rain.exp <- coef(MR_nls)[[1]]*exp(coef(MR_nls)[[2]]*NH3predict_m8$Rain.mean)

#
# Model
mR_lm_exp <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Rain.exp + offset(M1_fit), 
             weights=Weight, 
             data=NH3predict_m8)

# Model assumptions
# par(mfrow=c(1,3))
# qqnorm(resid(mR_lm_exp))
# qqline(resid(mR_lm_exp))
# plot(fitted(mR_lm_exp), NH3predict_m8$Weight*resid(mR_lm_exp))
# abline(h=0)
# hist(resid(mR_lm_exp))
# par(mfrow=c(1,1))

```

When non-linear rain is analysed separately, there is no effect of rain on NH3loss (F(`r anova(mR_lm_exp)$Df[1]`, `r anova(mR_lm_exp)$Df[2]`) =`r round(anova(mR_lm_exp)$F[1], 2)`, p `r ifelse(round(anova(mR_lm_exp)$P[1], 3)=="0", "<0.001", paste("=",round(anova(mR_lm_exp)$P[1], 3)))`). See the slope estimate from below:

```{r include=TRUE, echo=FALSE}

# Results
summary(mR_lm_exp)

```

\newpage

## Model A_Irrigation

```{r include=FALSE, echo=FALSE}

#########################################################################################################################################################
## MODEL A9

#
#
# Remove NA's from the dataset
NH3predict_m9 <- NH3predict %>% drop_na(irrigation)

```

Irrigation was modeled as a binomial variable due to heavy zero-inflation in the data.

```{r include=TRUE, echo=FALSE}
NH3predict_m9$irrigation <- as.numeric(as.character(NH3predict_m9$irrigation))
hist(NH3predict_m9$irrigation, breaks=50)

```

The number of observations in total in each category can be seen below:

```{r include=TRUE, echo=FALSE}

ddply(NH3predict_m9, "irrigation_cat", summarise, N=length(NH3loss))

```

```{r include=TRUE, echo=TRUE}
# Model 
m9_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ irrigation_cat + offset(M1_fit), 
            weights=Weight,
            data=NH3predict_m9)

```

The check for normality and heteroscedasticity of errors can be seen below:

```{r include=TRUE, echo=FALSE}
# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(m9_lm))
qqline(resid(m9_lm))
plot(fitted(m9_lm), NH3predict_m9$Weight*resid(m9_lm))
abline(h=0)
hist(resid(m9_lm))
par(mfrow=c(1,1))


```

There was no effect of irrigation on NH3-loss (F(`r anova(m9_lm)$Df[1]`, `r anova(m9_lm)$Df[2]`) =`r round(anova(m9_lm)$F[1], 2)`, p `r ifelse(round(anova(m9_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(m9_lm)$P[1], 3)))`).

```{r include=TRUE, echo=FALSE}
# Results
summary(m9_lm)

# Save fitted values for plotting
NH3predict_m9$M9_fits <- fitted(m9_lm)

```

```{r include=FALSE, echo=FALSE}

# Extracting correct R2-measure (including the effect of offset)

m9_lm_res <- lm(M1_res ~ irrigation_cat, 
            weights=Weight,
            data=NH3predict_m9)

round(summary(m9_lm_res)$adj.r.squared*100, 2)

```

Adjusted R2-value of the model was `r round(summary(m9_lm_res)$adj.r.squared, 5)`, and multiple R2 was `r round(summary(m9_lm_res)$r.squared,5)`. The negative adjusted R2-value indicates that the amount of variation accounted for by irrigation is negligible. In practice, it explain no variation in the data.

The marginal means for NH3 loss in no-irrigation and irrigation categories can be seen below.

```{r include=TRUE, echo=FALSE}

#
# Effect sizes
m9_est <- emmeans(m9_lm, ~irrigation_cat)
m9_est <- data.frame(m9_est)
m9_est

# Figure
ggplot(m9_est, aes(y=irrigation_cat, x=emmean, xmin=lower.CL, xmax=upper.CL)) +
  theme_classic(base_size = 11) +
  geom_point(size=3) + 
  geom_errorbarh(height=.1)  +
  geom_vline(xintercept=0, colour="black", linetype="dashed") +
  labs(x="Effect Size (NH3 loss (%))", y="Irrigation") +
  theme(legend.position = "none")

#
#
# 
```

\newpage

## Model A_Water input

Daily water input (mm) consists of irrigation and rainfall.

```{r include=TRUE, echo=FALSE}

#########################################################################################################################################################
## MODEL A10

#
#
# Remove NA's from the dataset
NH3predict_m10 <- NH3predict %>% drop_na(WaterPerDay)


```

```{r include=TRUE, echo=FALSE}
hist(NH3predict_m10$WaterPerDay, breaks=50)

```

```{r include=TRUE, echo=TRUE}
#  Model
m10_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ WaterPerDay + offset(M1_fit), 
             weights=Weight,
             data=NH3predict_m10)
```

Heteroscedasticity and normality of erros were as below:

```{r include=TRUE, echo=FALSE}
# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(m10_lm))
qqline(resid(m10_lm))
plot(fitted(m10_lm), NH3predict_m10$Weight*resid(m10_lm))
abline(h=0)
hist(resid(m10_lm))
par(mfrow=c(1,1))
```

There was no effect of water input on NH3-loss (F(`r anova(m10_lm)$Df[1]`, `r anova(m10_lm)$Df[2]`) =`r round(anova(m10_lm)$F[1], 2)`, p `r ifelse(round(anova(m10_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(m10_lm)$P[1], 3)))`).

```{r include=TRUE, echo=FALSE}

# Results
summary(m10_lm)
anova(m10_lm)

# Save fitted values for plotting
NH3predict_m10$M10_fits <- fitted(m10_lm)

```

```{r include=FALSE, echo=FALSE}

# Extracting correct R2-measure (including the effect of offset)

m10_lm_res <- lm(M1_res ~ WaterPerDay, 
             weights=Weight,
             data=NH3predict_m10)

round(summary(m10_lm_res)$adj.r.squared*100, 2)

```

Adjusted R2-value of the model was `r round(summary(m10_lm_res)$adj.r.squared, 5)`, and multiple R2 was `r round(summary(m10_lm_res)$r.squared,5)`. The negative adjusted R2-value indicates that the amount of variation accounted for by daily water input is negligible. In practice, it explain no variation in the data.

Figure below shows a linear approximation of the effect of daily water input (mm) on NH3-emissions (linear slope back-transformed from the square-root scale).

```{r include=TRUE, echo=FALSE, message=FALSE}

# Figure
ggplot(NH3predict_m10, aes(y=NH3loss, x=WaterPerDay)) + 
  theme_classic(base_size = 11) +
  geom_point(colour="gray75") +
  geom_smooth(method="lm", se=T, colour="black", aes(y=(M10_fits^2)-4)) +
  labs(x="Water input (mm)", y = "NH3 loss (%)") 

#
#
# 

```

\newpage

## Model A_Experiment duration

Duration of the experiment (days).

```{r include=TRUE, echo=FALSE}

#########################################################################################################################################################
## MODEL A10

#
#
# Remove NA's from the dataset
NH3predict_m11 <- NH3predict[which(is.na(NH3predict$Duration)==FALSE),]
NH3predict_m11b <- NH3predict_m11[which(is.na(NH3predict$AirTemperature)==FALSE),]
summary(NH3predict_m11$Duration)
```

```{r include=TRUE, echo=FALSE}
hist(NH3predict_m11$Duration, breaks=50)

```

```{r include=TRUE, echo=TRUE}
#  Model
m11_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Duration + offset(M1_fit), 
             weights=Weight,
             data=NH3predict_m11)
```

Heteroscedasticity and normality of erros were as below:

```{r include=TRUE, echo=FALSE, warning=FALSE}
# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(m11_lm))
qqline(resid(m11_lm))
plot(fitted(m11_lm), NH3predict_m10$Weight*resid(m11_lm))
abline(h=0)
hist(resid(m11_lm))
par(mfrow=c(1,1))
```

There was no effect of duration of the experiment on NH3-loss (F(`r anova(m11_lm)$Df[1]`, `r anova(m11_lm)$Df[2]`) =`r round(anova(m11_lm)$F[1], 2)`, p `r ifelse(round(anova(m11_lm)$P[1], 3)=="0", "<0.001", paste("=",round(anova(m11_lm)$P[1], 3)))`).

```{r include=TRUE, echo=FALSE}

# Results
summary(m11_lm)
anova(m11_lm)

# Save fitted values for plotting
NH3predict_m11$M11_fits <- fitted(m11_lm)

```

```{r include=FALSE, echo=FALSE}

# Extracting correct R2-measure (including the effect of offset)

m11_lm_res <- lm(M1_res ~ Duration, 
             weights=Weight,
             data=NH3predict_m11)

round(summary(m11_lm_res)$adj.r.squared*100, 2)

```

Adjusted R2-value of the model was `r round(summary(m11_lm_res)$adj.r.squared, 5)`, and multiple R2 was `r round(summary(m11_lm_res)$r.squared,5)`. The negative adjusted R2-value indicates that the amount of variation accounted for by duration of the experiment is negligible. In practice, it explain no variation in the data.

Figure below shows a linear approximation of the effect of daily water input (mm) on NH3-emissions (linear slope back-transformed from the square-root scale).

```{r include=TRUE, echo=FALSE, message=FALSE}

# Figure
ggplot(NH3predict_m11, aes(y=NH3loss, x=Duration)) + 
  theme_classic(base_size = 11) +
  geom_point(colour="gray75") +
  geom_smooth(method="lm", se=T, colour="black", aes(y=(M11_fits^2)-4)) +
  labs(x="Experiment duration (days)", y = "NH3 loss (%)") 

#
#
# 
# Remove NA's from the dataset
# NH3predict_m12 <- NH3predict[which(is.na(NH3predict$AEZ)==FALSE),]
# summary(NH3predict_m12$AEZ)
# #  Model
# m12_lm <- lm(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ AEZ + Crop + Crop * AEZ + offset(M1_fit), 
#             weights=Weight,
#            data=NH3predict_m12)
# summary(m12_lm)
# anova(m12_lm)

```

\newpage

## Combined model with all variables of effect

### Model LO1_Base model and pH

```{r include=TRUE, echo=FALSE}


NH3predict_mALL <- NH3predict[which(is.na(NH3predict$pH_cat)==FALSE),]

mALL1_lmer <- lmer(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Fertiliser + Application + Method +  pH_cat + Fertiliser:pH_cat +(1|Ref),
                   weights=Weight,
                   control=lmerControl(optimizer="bobyqa"),
                   data=NH3predict_mALL)
anova(mALL1_lmer)
summary(mALL1_lmer )

r.squaredGLMM(mALL1_lmer)

```

Total observations = `r length(NH3predict_mALL$NH3loss)` \normalsize

Model LO1 explains `r round(r.squaredGLMM(mALL1_lmer)[2]*100, 2)` % of the variation in the data (conditional R2), with the fixed effects explaining `r round(r.squaredGLMM(mALL1_lmer)[1]*100, 2)` % (marginal R2). The R2-values for linear mixed effects model were calculated according to Nagakawa & Schielzeth (2013).

```{r include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#
##### 
# Estimates for each main effect


#
# Fertiliser

# Effect sizes
at <- list("Application"="Broadcast", "Method"="Micromet", "pH_cat"="normal")
mALL1_est_F <- emmeans(mALL1_lmer, ~ Fertiliser, type="response", at=at)
mALL1_est_F  <- data.frame(mALL1_est_F)

#
# Application

# Effect sizes
at <- list("Fertiliser"="urea+", "Method"="Micromet", "pH_cat"="normal")
mALL1_est_A <- emmeans(mALL1_lmer, ~ Application, type="response", at=at)
mALL1_est_A <- data.frame(mALL1_est_A)

#
# Method

# Effect sizes
at <- list("Fertiliser"="urea+", "Application"="Broadcast", "pH_cat"="normal")
mALL1_est_M <- emmeans(mALL1_lmer, ~ Method, type="response", at=at)
mALL1_est_M <- data.frame(mALL1_est_M)

#
# Soil pH

# Effect size
at <- list("Fertiliser"="urea+", "Application"="Broadcast", "Method"="Micromet")
mALL1_est_pH <- emmeans(mALL1_lmer, ~ pH_cat, type="response", at=at)
mALL1_est_pH <- data.frame(mALL1_est_pH)

#
# Interaction

# Effect size
at <- list("Application"="Broadcast", "Method"="Micromet", "Place"="Outdoor")
mALL1_est_pHFe <- emmeans(mALL1_lmer, pairwise ~ pH_cat | Fertiliser, regrid="response", at=at)
mALL1_est_pHFe <- data.frame(mALL1_est_pHFe)


#

## 
# Collating a dataframe with all effect sizes

# Vectors for emmeans-objects and fectors
Model_vec <- c("mALL1_est_F", "mALL1_est_A", "mALL1_est_M", "mALL1_est_pH")
Factor_vec <- c("Fertiliser", "Application", "Method", "Soil pH")

# Creating an empty dataframe
mALL1_estimates <- data.frame(Factor=factor(), Level=factor(), Effect=numeric(), SE=numeric(), df=numeric(), lower.CL=numeric(), upper.CL=numeric())

# Loop for extracting all estimates
for(i in c(1:4)){ # Cycling through models
  m_est <- data.frame(get(Model_vec[i]))
  mALL1_estimates <- rbind(mALL1_estimates, data.frame(Factor=Factor_vec[i],
                                                 Level=m_est[,1],
                                                 Effect=m_est[,2],
                                                 SE=m_est[,3],
                                                 df=m_est[,4],
                                                 lower.CL=m_est[,5],
                                                 upper.CL=m_est[,6]))
}


mALL1_est_pHFe$Level <- paste(mALL1_est_pHFe$Fertiliser, ":", mALL1_est_pHFe$pH_cat)


# Add the interaction
mALL1_estimates <- rbind(mALL1_estimates, data.frame(Factor="Fertiliser:Soil pH",
                                                 Level=filter(mALL1_est_pHFe, contrast==".")$Level,
                                                 Effect=filter(mALL1_est_pHFe, contrast==".")$response,
                                                 SE=filter(mALL1_est_pHFe, contrast==".")$SE,
                                                 df=filter(mALL1_est_pHFe, contrast==".")$df,
                                                 lower.CL=filter(mALL1_est_pHFe, contrast==".")$lower.CL,
                                                 upper.CL=filter(mALL1_est_pHFe, contrast==".")$upper.CL))

mALL1_estimates

#write.csv(mALL1_estimates, "C:/Users/au721810/Projects/2022, NH3 loss, Hutchings/Analysis/2022_NH3loss_Hutchings/output/Model_estimates/2022_NH3loss_Model1_EffectsM13.csv")

```

### Model LO2_Base model, pH, and Application rate


```{r include=TRUE, echo=FALSE}

NH3predict_mALL2 <- NH3predict[which(is.na(NH3predict$pH_cat)==FALSE & is.na(NH3predict$Application.rate)==FALSE),]

mALL2_lmer <- lmer(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Fertiliser + Application + Method  +                     pH_cat + Fertiliser:pH_cat + Application.rate + (1|Ref),
                   weights=Weight,
                   control=lmerControl(optimizer="bobyqa"),
                   data=NH3predict_mALL2)
anova(mALL2_lmer)
summary(mALL2_lmer)
r.squaredGLMM(mALL2_lmer)
```

\normalsize

Model LO2 explains `r round(r.squaredGLMM(mALL2_lmer)[2]*100, 2)` % of the variation in the data (conditional R2), with the fixed effects explaining `r round(r.squaredGLMM(mALL2_lmer)[1]*100, 2)` % (marginal R2). Total observations = `r length(NH3predict_mALL2$NH3loss)`

### Model LO3_Base model, pH and temperature

```{r include=TRUE, echo=FALSE}


NH3predict_mALL3 <- NH3predict[which(is.na(NH3predict$pH_cat)==FALSE &  is.na(NH3predict$AirTemperature)==FALSE),]

mALL3_lmer <- lmer(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Fertiliser + Application + Method  +  pH_cat + Fertiliser:pH_cat + AirTemperature +(1|Ref),
                   weights=Weight,
                   control=lmerControl(optimizer="bobyqa"),
                   data=NH3predict_mALL3)
anova(mALL3_lmer)
summary(mALL3_lmer )
r.squaredGLMM(mALL3_lmer)

```

Total observations = `r length(NH3predict_mALL3$NH3loss)`

Model LO3 explains `r round(r.squaredGLMM(mALL3_lmer)[2]*100, 2)` % of the variation in the data (conditional R2), with the fixed effects explaining `r round(r.squaredGLMM(mALL3_lmer)[1]*100, 2)` % (marginal R2). The R2-values for linear mixed effects model were calculated according to Nagakawa & Schielzeth (2013).


### Model LO4_Base variables, pH, application rate, and temperature

```{r include=TRUE, echo=FALSE}
#
# Remove NA's from the dataset
NH3predict_mALL4 <- NH3predict[which(is.na(NH3predict$pH_cat)==FALSE & is.na(NH3predict$Application.rate)==FALSE & is.na(NH3predict$AirTemperature)==FALSE),]

```

```{r include=TRUE, echo=TRUE}

mALL4_lmer <- lmer(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Fertiliser + Application + Method  + 
                    pH_cat + Fertiliser:pH_cat + 
                    Application.rate + AirTemperature +
                    (1|Ref),
                   weights=Weight,
                   control=lmerControl(optimizer="bobyqa"),
                   data=NH3predict_mALL4)
```

Total observations = `r length(NH3predict_mALL4$NH3loss)`

```{r include=TRUE, echo=FALSE}

# Model assumptions
par(mfrow=c(1,3))
qqnorm(resid(mALL4_lmer))
qqline(resid(mALL4_lmer))
plot(fitted(mALL4_lmer), NH3predict_mALL4$Weight*resid(mALL4_lmer))
abline(h=0)
hist(resid(mALL4_lmer))
par(mfrow=c(1,1))


# Variance inflation 
#vif(mALL4_lmer)

# Autocorrelation
#acf(resid(mALL4_lmer), type="correlation")

```

```{r include=TRUE, echo=FALSE}

anova(mALL4_lmer)
summary(mALL4_lmer)
r.squaredGLMM(mALL4_lmer)

```

\normalsize

Model LO4 explains `r round(r.squaredGLMM(mALL4_lmer)[2]*100, 2)` % of the variation in the data (conditional R2), with the fixed effects explaining `r round(r.squaredGLMM(mALL4_lmer)[1]*100, 2)` % (marginal R2).

```{r include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#
##### 
# Estimates for each main effect


#
# Fertiliser

# Effect sizes
at <- list("Application"="Broadcast", "Method"="Micromet", "pH_cat"="normal")
mALL4_est_F <- emmeans(mALL4_lmer, ~ Fertiliser, type="response", at=at)
mALL4_est_F  <- data.frame(mALL4_est_F)

#
# Application

# Effect sizes
at <- list("Fertiliser"="urea+", "Method"="Micromet", "pH_cat"="normal")
mALL4_est_A <- emmeans(mALL4_lmer, ~ Application, type="response", at=at)
mALL4_est_A <- data.frame(mALL4_est_A)

#
# Method

# Effect sizes
at <- list("Fertiliser"="urea+", "Application"="Broadcast", "pH_cat"="normal")
mALL4_est_M <- emmeans(mALL4_lmer, ~ Method, type="response", at=at)
mALL4_est_M <- data.frame(mALL4_est_M)

#
# Soil pH

# Effect size
at <- list("Fertiliser"="urea+", "Application"="Broadcast", "Method"="Micromet")
mALL4_est_pH <- emmeans(mALL4_lmer, ~ pH_cat, type="response", at=at)
mALL4_est_pH <- data.frame(mALL4_est_pH)

#
# Interaction

# Effect size
at <- list("Application"="Broadcast", "Method"="Micromet")
mALL4_est_pHFe <- emmeans(mALL4_lmer, pairwise ~ pH_cat | Fertiliser, regrid="response", at=at)
mALL4_est_pHFe <- data.frame(mALL4_est_pHFe)


#

## 
# Collating a dataframe with all effect sizes

# Vectors for emmeans-objects and fectors
Model_vec <- c("mALL4_est_F", "mALL4_est_A", "mALL4_est_M", "mALL4_est_pH")
Factor_vec <- c("Fertiliser", "Application", "Method", "Soil pH")

# Creating an empty dataframe
mALL4_estimates <- data.frame(Factor=factor(), Level=factor(), Effect=numeric(), SE=numeric(), df=numeric(), lower.CL=numeric(), upper.CL=numeric())

# Loop for extracting all estimates
for(i in c(1:4)){ # Cycling through models
  m_est <- data.frame(get(Model_vec[i]))
  mALL4_estimates <- rbind(mALL4_estimates, data.frame(Factor=Factor_vec[i],
                                                 Level=m_est[,1],
                                                 Effect=m_est[,2],
                                                 SE=m_est[,3],
                                                 df=m_est[,4],
                                                 lower.CL=m_est[,5],
                                                 upper.CL=m_est[,6]))
}


mALL4_est_pHFe$Level <- paste(mALL4_est_pHFe$Fertiliser, ":", mALL4_est_pHFe$pH_cat)


# Add the interaction
mALL4_estimates <- rbind(mALL4_estimates, data.frame(Factor="Fertiliser:Soil pH",
                                                 Level=filter(mALL4_est_pHFe, contrast==".")$Level,
                                                 Effect=filter(mALL4_est_pHFe, contrast==".")$response,
                                                 SE=filter(mALL4_est_pHFe, contrast==".")$SE,
                                                 df=filter(mALL4_est_pHFe, contrast==".")$df,
                                                 lower.CL=filter(mALL4_est_pHFe, contrast==".")$lower.CL,
                                                 upper.CL=filter(mALL4_est_pHFe, contrast==".")$upper.CL))

mALL4_estimates

#write.csv(mALL_estimates, "C:/Users/au721810/Projects/2022, NH3 loss, Hutchings/Analysis/2022_NH3loss_Hutchings/output/Model_estimates/2022_NH3loss_Model1_EffectsMALL.csv")

```

The same effect sizes (and their 95% confidence interval) are shown below:

```{r include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

# Figure
ggplot(mALL4_estimates, aes(y=Level, x=Effect, xmin=lower.CL, xmax=upper.CL, colour=Factor)) +
  theme_classic(base_size = 11) +
  facet_grid(Factor ~ ., scales = "free_y", space = "free_y") +
  geom_point(size=3) + 
  geom_errorbarh(height=.1)  +
  geom_vline(xintercept=0, colour="black", linetype="dashed") +
  scale_color_brewer(palette = "Dark2") +
  labs(x="Effect Size (NH3 loss (%))", y="") +
  theme(legend.position = "none")


```


### Model LO5_Base model, pH, temperature, rainfall and Application rate


```{r include=TRUE, echo=FALSE}

NH3predict_mALL5 <- NH3predict[which(is.na(NH3predict$pH_cat)==FALSE & 
is.na(NH3predict$Application.rate)==FALSE &                   is.na(NH3predict$AirTemperature)==FALSE & is.na(NH3predict$Rain.mean)==FALSE),]

mALL5_lmer <- lmer(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Fertiliser + Application + Method  +  pH_cat + Fertiliser:pH_cat + Application.rate + AirTemperature + Rain.mean + AirTemperature * Rain.mean + (1|Ref),
                   weights=Weight,
                   control=lmerControl(optimizer="bobyqa"),
                   data=NH3predict_mALL5)
anova(mALL5_lmer)
summary(mALL5_lmer)
r.squaredGLMM(mALL5_lmer)
```

\normalsize

Model LO5 explains `r round(r.squaredGLMM(mALL5_lmer)[2]*100, 2)` % of the variation in the data (conditional R2), with the fixed effects explaining `r round(r.squaredGLMM(mALL5_lmer)[1]*100, 2)` % (marginal R2). Total observations = `r length(NH3predict_mALL5$NH3loss)`

### Model LO6_Base model, pH, temperature, non-linear rainfall and Application rate


```{r include=TRUE, echo=FALSE}

NH3predict_mALL6 <- NH3predict[which(is.na(NH3predict$SoilpH)==FALSE &
is.na(NH3predict$Application.rate)==FALSE &                   is.na(NH3predict$AirTemperature)==FALSE & is.na(NH3predict$Rainfall)==FALSE),]

NH3predict_mALL6$pH_cat <- factor(ifelse(NH3predict_mALL6$SoilpH<=7, "normal", "chalky"))
NH3predict_mALL6$Rain.mean <- NH3predict_mALL6$Rainfall/NH3predict_mALL6$Duration

NH3predict_mALL6$Rain.exp <- coef(MR_nls)[[1]]*exp(coef(MR_nls)[[2]]*NH3predict_mALL6$Rain.mean)
#NH3predict_mALL6$Weight <- 1/lm(abs(resid(model)) ~ #fitted(model))$fitted.values^2

mALL6_lmer <- lmer(sign(NH3loss) * abs(NH3loss) ^ (1/3) ~ Fertiliser + Application + Method  +  pH_cat + Fertiliser:pH_cat + Application.rate + AirTemperature + Rain.exp + AirTemperature * Rain.exp + (1|Ref),
                   weights=Weight,
                   control=lmerControl(optimizer="bobyqa"),
                   data=NH3predict_mALL6)
anova(mALL6_lmer)
summary(mALL6_lmer)
```

\normalsize

Model LO6 explains `r round(r.squaredGLMM(mALL6_lmer)[2]*100, 2)` % of the variation in the data (conditional R2), with the fixed effects explaining `r round(r.squaredGLMM(mALL6_lmer)[1]*100, 2)` % (marginal R2). Total observations = `r length(NH3predict_mALL6$NH3loss)`


# GLMM

```{r include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

mALL_glmm <- glmmTMB(NH3loss ~ Fertiliser + Application + Method + 
                    pH_cat + Fertiliser:pH_cat + 
                    Clay + AirTemperature +
                    (1|Ref),
                   dispformula= ~ Fertiliser + Application + Method + pH_cat,
                   family=tweedie(link="log"),
                   data=NH3predict_mALL)

car::Anova(mALL_glmm)
summary(mALL_glmm)

```

```{r include=T, echo=F, message=F, warning=F}

plot(simulateResiduals(mALL_glmm))

```

Total observations = `r length(NH3predict_mALL$NH3loss)` \normalsize

Model LO_GLIMM explains `r round(r.squaredGLMM(mALL_glmm)[2]*100, 2)` % of the variation in the data (conditional R2), with the fixed effects explaining `r round(r.squaredGLMM(mALL_glmm)[1]*100, 2)` % (marginal R2). The R2-values for linear mixed effects model were calculated according to Nagakawa & Schielzeth (2013).

```{r include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#
##### 
# Estimates for each main effect


#
# Fertiliser

# Effect sizes
at <- list("Application"="Broadcast", "Method"="Micromet", "pH_cat"="normal")
mALLglmm_est_F <- emmeans(mALL_glmm, ~ Fertiliser, type="response", at=at)
mALLglmm_est_F  <- data.frame(mALLglmm_est_F)

#
# Application

# Effect sizes
at <- list("Fertiliser"="urea+", "Method"="Micromet", "pH_cat"="normal")
mALLglmm_est_A <- emmeans(mALL_glmm, ~ Application, type="response", at=at)
mALLglmm_est_A <- data.frame(mALLglmm_est_A)

#
# Method

# Effect sizes
at <- list("Fertiliser"="urea+", "Application"="Broadcast", "Place"="Outdoor", "pH_cat"="normal")
mALLglmm_est_M <- emmeans(mALL_glmm, ~ Method, type="response", at=at)
mALLglmm_est_M <- data.frame(mALLglmm_est_M)

#
# Soil pH

# Effect size
at <- list("Fertiliser"="urea+", "Application"="Broadcast", "Method"="Micromet")
mALLglmm_est_pH <- emmeans(mALL_glmm, ~ pH_cat, type="response", at=at)
mALLglmm_est_pH <- data.frame(mALLglmm_est_pH)

#
# Interaction

# Effect size
at <- list("Application"="Broadcast", "Method"="Micromet")
mALLglmm_est_pHFe <- emmeans(mALL_glmm, pairwise ~ pH_cat | Fertiliser, regrid="response", at=at)
mALLglmm_est_pHFe <- data.frame(mALLglmm_est_pHFe)


#

## 
# Collating a dataframe with all effect sizes

# Vectors for emmeans-objects and fectors
Model_vec <- c("mALLglmm_est_F", "mALLglmm_est_A", "mALLglmm_est_M", "mALLglmm_est_pH")
Factor_vec <- c("Fertiliser", "Application", "Method", "Soil pH")

# Creating an empty dataframe
mALLglmm_estimates <- data.frame(Factor=factor(), Level=factor(), Effect=numeric(), SE=numeric(), df=numeric(), lower.CL=numeric(), upper.CL=numeric())

# Loop for extracting all estimates
for(i in c(1:4)){ # Cycling through models
  m_est <- data.frame(get(Model_vec[i]))
  mALLglmm_estimates <- rbind(mALLglmm_estimates, data.frame(Factor=Factor_vec[i],
                                                 Level=m_est[,1],
                                                 Effect=m_est[,2],
                                                 SE=m_est[,3],
                                                 df=m_est[,4],
                                                 lower.CL=m_est[,5],
                                                 upper.CL=m_est[,6]))
}


mALLglmm_est_pHFe$Level <- paste(mALLglmm_est_pHFe$Fertiliser, ":", mALLglmm_est_pHFe$pH_cat)


# Add the interaction
mALLglmm_estimates <- rbind(mALLglmm_estimates, data.frame(Factor="Fertiliser:Soil pH",
                                                 Level=filter(mALLglmm_est_pHFe, contrast==".")$Level,
                                                 Effect=filter(mALLglmm_est_pHFe, contrast==".")$response,
                                                 SE=filter(mALLglmm_est_pHFe, contrast==".")$SE,
                                                 df=filter(mALLglmm_est_pHFe, contrast==".")$df,
                                                 lower.CL=filter(mALLglmm_est_pHFe, contrast==".")$asymp.LCL,
                                                 upper.CL=filter(mALLglmm_est_pHFe, contrast==".")$asymp.UCL))

mALLglmm_estimates


# Figure
ggplot(mALLglmm_estimates, aes(y=Level, x=Effect, xmin=lower.CL, xmax=upper.CL, colour=Factor)) +
  theme_classic(base_size = 11) +
  facet_grid(Factor ~ ., scales = "free_y", space = "free_y") +
  geom_point(size=3) + 
  geom_errorbarh(height=.1)  +
  geom_vline(xintercept=0, colour="black", linetype="dashed") +
  scale_color_brewer(palette = "Dark2") +
  labs(x="Effect Size (NH3 loss (%))", y="") +
  theme(legend.position = "none")


```

# References

Astivia, Oscar L. Olvera and Zumbo, Bruno D. (2019) "Heteroskedasticity in Multiple Regression Analysis: What it is, How to Detect it and How to Solve it with Applications in R and SPSS," Practical Assessment, Research, and Evaluation: Vol. 24 , Article 1. DOI: <https://doi.org/10.7275/q5xr-fr95>

Douglas Bates, Martin Maechler, Ben Bolker, Steve Walker (2015). Fitting Linear Mixed-Effects Models Using lme4. Journal of Statistical Software, 67(1), 1-48. [https://doi:10.18637/jss.v067.i01](https://doi:10.18637/jss.v067.i01){.uri}

Nakagawa, S, Schielzeth, H. (2013). A general and simple method for obtaining R\^2 from Generalized Linear Mixed-effects Models. Methods in Ecology and Evolution 4: 133-142. <https://doi.org/10.1111/j.2041-210x.2012.00261.x>
